Supplemental Methods

### 2024-03-29 script/CoverLetter.Rmd
appendix remade, similar to last...

---
output: pdf_document
---

\begin{flushleft}
Dr. Aaron Ellison \newline
Executive Editor \newline
Methods in Ecology and Evolution \newline
Senior Research Fellow in Ecology \textit{Emeritus} \newline
Harvard Forest \newline
\end{flushleft}

March 17^th^, 2024

Attached is our standard article submission "*Protocol to increase accuracy and fidelity of pollen meta-genomic barcoding using Angiosperms353: a case study using pollen loads from wild bumble bees*".

In this paper we develop an intuitive approach to generate plant reference sequence databases tailored to specific areas for metabarcoding projects, as well as a lens to critically evaluate metagenomic results using phenological data.
This spatio-temporal approach utilizes, among other sources - museum and citizen science records, species distribution modelling, and the weibull distribution, to create baseline environmental data in an approach which is globally applicable.
Further we use bait-capture, via the popular Angiosperms 353 probes, as our method to generate sequence reads for metabarcoding.

Given the rising popularity of metabarcoding and metagenomics, especially for projects over large spatial domains and with multiple seasonal time points, we believe that our method provides a framework for meaningfully assessing and contextualizing results. 
We have observed that many studies utilizing metabacoding generate spurious results, which are precluded due to spatial or temporal incongruities with the focal animal species and plants. 
However, a simple and intuitive framework for programmatically assessing and rectifying these issues are lacking.

We believe that the use of the Angiosperms353 probes was novel at the inception of the study, and are aware of several manuscripts utilizing them which will be forthcoming in a variety of journals.

Our submission contains X supplementary materials, X figures and X tables. It unfortunately is currently at roughly 8,900 words. 
However *circa* 900 words are dedicated to a plain text overview of the methods, a section which we believe elucidates the integration of the various approaches, and clarifies adoption by subsequent users. 
Considerable length is dedicated to the methods, which contain five years of field work, and similarly thorough microscopy, and molecular work; but which we have cut by nearly 1000 words.
The full details or our methods are provided as supplementary materials.

We have no conflicts of interest to disclose. Additionally, this work - nor any part of it - has been submitted to any other journal. Although, it is cited and discussed in our Grimes Review which is currently in review for *Journal of Ecology*.

Please address all correspondence regarding this manuscript to me at rbenkendorf@chicagobotanic.org

Thank you for considering this work.

\begin{flushleft}
Reed Clark Benkendorf \newline
Senior (Field) Botanist \& Senior Spatial Data Specialist  \newline
The Chicago Botanic Garden  \newline
rbenkendorf@chicagobotanic.org
\end{flushleft}

\begin{flushleft}
Jeremie B. Fant, Ph.D \newline
Medard and Elizabeth Welch Director of Conservation \newline
The Chicago Botanic Garden  \newline
jfant@chicagobotanic.org
\end{flushleft}

>>>>

---
output:
  pdf_document: default
  word_document: default
---

\begin{flushleft}
Dr. Aaron Ellison \newline
Executive Editor \newline
Methods in Ecology and Evolution \newline
\end{flushleft}

\hfill March 17^th^, 2024

Dear Dr. Ellison,

We are writing to propose an original research article entitled  "*Protocol to increase accuracy and fidelity of pollen meta-genomic barcoding using Angiosperms353: a case study using pollen loads from wild bumble bees*".

Many meta-barcoding studies that include plant material are limited due to lack of reliable universal barcodes and the extensive potential candidate list (~350,000 plant species) that requires extensive computational and resource to identify best matches.
In this paper we develop an intuitive approach to generate plant reference sequence databases tailored to specific areas for metabarcoding projects, as well as a lens to critically evaluate metagenomic results using phenological data. 
This spatio-temporal approach utilizes, among other sources - museum and citizen science records, species distribution modelling, and the weibull distribution -  to create baseline environmental data in an approach which is globally applicable

Further, we use bait-capture, via the popular Angiosperms 353 probes, to generate multiple sequence reads to increase the accuracy of species identification.
We believe that the use of the Angiosperms353 probes was novel at the inception of the study, and are aware of several manuscripts utilizing them which will be forthcoming in a variety of journals.

Finally, many studies utilizing metabacoding generate spurious results, which are spatially or temporally incongruities with the focal animal species and plants/ There are few simple and intuitive framework for programmatically assessing and rectifying these issues are lacking. 
In this study, we also provide a framework for meaningfully assessing and contextualizing the meta-barcoding results.

We have no conflicts of interest to disclose.

### 2024-03-29 script/StatementOnInclusion.Rmd
appendix remade, similar to last...

---
title: "StatementOnInclusion"
output: pdf_document
---

'Statement on inclusion'

>>>>

---
title: "Statement On Inclusion"
output:
  word_document: default
  pdf_document: default
---

Our studies focal area was within the Southern Rockies Mountains of Western North America and drew on natural history data collected throughout the region. 
Two of our authors are *staff?affiliates?* at the research station which novel data was collected at. 
Prior to matriculation our corresponding author worked at the largest institute for applied ecology in the Rocky Mountains, and has since worked as a contractor for the largest land management agency in the United States and is both based in and works in this geographic area.

### 2024-03-29 script/SupplementaryInformation.Rmd
appendix remade, similar to last...

---
title: "Supplemental Information"
author: "steppe"
output: pdf_document
---

@@

# Site Info

@@

# Table S2 - Models used for Species Distribution Model Ensembles

@@

Table S3 - Species Distribution Models Predictors

@@

Appendix 3 - Molecular Reference Specimen Table

@@

Appendix 6 - Pollen Key

@@

Appendix 9 - All Species in the Sequence Databases

@@

Appendix 11 - Reads Per Loci

@@

```{r READS PER LOCI - METAGENOMICS, echo = F, out.width = '98%'}
knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
```

@@

Appendix 12 - Comparison of Kraken2, Bracken, and BLAST

@@

>>>>

---
title: "Supplemental Information"
author: "steppe"
output:
  pdf_document: default
---

@@

Figure 1 - Site Info

@@

Table 2 - Models used for Species Distribution Model Ensembles

@@

Table 3 - Species Distribution Models Predictors

@@

Table 4 - Molecular Reference Specimen Table

@@

Figure 2 - Pollen Key

@@

Table 5 - All Species in the Sequence Databases

@@

Figure 3 - Reads Per Loci

@@

@@

Figure 4 - Percent Loci Matched by Sample 
```{r READS PER LOCI - METAGENOMICS, echo = F, out.width = '98%'}
knitr::include_graphics('../graphics/plots/Percent_loci_matched_by_sample.png')
```

Figure 5 - Comparison of Kraken2, Bracken, and BLAST

@@

Figure 6 - Microscopy Rarefaction, Abundance

```{r, echo = F, out.width="90%"}
knitr::include_graphics('../graphics/plots/SppAbundance.png')
```

Figure 7 - Microscopy Rarefaction, Richness

```{r, echo = F, out.width="90%"}
knitr::include_graphics('../graphics/plots/species_richness_rarefaction.png')
```

Table 6 - Binary reclassification process

```{r, echo = F, out.width="90%"}
knitr::include_graphics('../graphics/tables/reclassify_reads_table.pdf')
```

### 2024-03-29 script/methods_supplemental.Rmd
appendix remade, similar to last...

>>>>

### Comparision of Sequence classifications pre and post processing

Using wilcox_effsize, with a one-sided hypothesis of greater, we have strong evidence of (p < 1e-04, wilcoxsign_test) an effect size of 0.732 95% CI [0.57, 0.84, n = 40, bootstrap replicates = 1000] (@kassambara2023rstatix, @Hothorn2006coin).
